---
bibliography: references.bib
---

# Overall Objectives and Research Questions

## Primary Objective

To develop, validate, and implement PK/PD models derived from the published literature into clinically usable MIPD tools.

## Secondary Objectives

-   Conduct systematic and repeatable literature reviews of population PK/PD models for selected antimicrobial and antiviral agents.
-   Reconstruct and harmonize published models into a reproducible computational framework.
-   Perform external validation using real-world clinical data (e.g., TDM datasets) to assess predictive performance and clinical plausibility.
-   Develop Shiny-based decision support tools implementing the validated models for simulation, Bayesian forecasting, and target attainment evaluation.
-   Establish a framework for periodic evidence updating, enabling the inclusion of newly published models and emerging agents (e.g., rezafungin, cabotegravir).
-   Develop a formal clinical trial protocol for the implementation of MIPD in clinical practice, including the design of a RCT to evaluate clinical, pharmacological, and patient-centered outcomes of MIPD-guided therapy.

### Literature Search Strategy

Systematic searches will be performed in PubMed/MEDLINE and Embase to identify parametric population pharmacokinetic (popPK) models for the selected agents. Searches will be adapted to each database’s syntax and controlled vocabulary when applicable. The strategy will focus on retrieving studies that report parametric nonlinear mixed-effects models.

The methodological search block described below will be applied consistently across all antibiotics.

***Pubmed***

```{r}
library(glue)
glue::glue(  "
  population pharmacokinetic*[Title/Abstract]
  OR
  (pharmacokinetic*[Title/Abstract] AND population*[Title/Abstract])
  OR
  popPK[Title/Abstract]
  OR
  NONMEM[Title/Abstract]
  OR
  NLME[Title/Abstract])
  OR
  nonlinear mixedeffect
  OR
  non linear mixed effect
             "
)
```

***Embase***

```{r}
glue::glue("
  population pharmacokinetic*.ti,ab.
  OR
  (pharmacokinetic*.ti,ab. AND population*.ti,ab.)
  OR
  popPK.ti,ab.
  OR
  NONMEM.ti,ab.
  OR
  NLME.ti,ab.
  OR
  \"nonlinear mixed effect*\".ti,ab.
  OR
  \"non linear mixed effect*\".ti,ab.
  "
)
```

This methodological block will be combined with drug-specific terms to generate individual search strategies for each antibiotic of interest, including:

-   Dalbavancin
-   Vancomycin
-   Piperacillin/tazobactam
-   Ceftazidime
-   Cefepime
-   Meropenem

Each antibiotic will therefore be associated with its own dedicated search. For piperacillin/tazobactam, both models describing piperacillin alone and models jointly describing piperacillin and tazobactam will be considered eligible for inclusion. In particular, population PK models that simultaneously characterize the pharmacokinetics of both compounds, or that explicitly model their interaction through shared covariates or correlated random effects, will be included where available.

In addition to database searches, citation-chasing techniques will be employed to identify relevant studies that may not be captured through electronic searches alone. This process will include backward citation chasing, by screening the reference lists of all included articles and relevant reviews, as well as forward citation chasing, by identifying subsequent publications that have cited the included studies using citation indexes.

### Inclusion and Exclusion Criteria

Studies will be included if they meet the following criteria: - Report a parametric population pharmacokinetic model developed using nonlinear mixed-effects modeling techniques. - Provide sufficient model details to allow for reconstruction, including structural model definitions, parameter estimates, interindividual variability terms, residual error models, and covariate relationships. - Published in peer-reviewed journals. - Written in English. Studies will be excluded if they: - Describe non-parametric or semi-parametric population PK models. - Lack sufficient detail for model reconstruction. - Are conference abstracts, reviews, editorials, or non-peer-reviewed articles.

### Model Construction

For each antibiotic identified through the systematic literature review, published parametric population PK/PD models will be reconstructed using a standardized and reproducible computational workflow. Model reconstruction will be based exclusively on information reported in the original publications, including structural model definitions, parameter estimates, interindividual variability terms, residual error models, and covariate relationships.

Published models will be translated into a common modeling framework using R-based tools. Here is an example of a published dalbavancin model [@cojutti2022] structure written in `rxode2` syntax [@wang2023].

The model is a two-compartment parametric model with first-order elimination, incorporating covariate effect of estimated glomerular filtration rate, calculated using the CKD-EPI formula. It was developed using data from 15 critically ill adult patients with bloodstream infections or endocarditis, with serial plasma samples collected up to 21 days post-infusion.

```{r}
#| echo: true
#| eval: false

 Cojutti2022 <- list(
  ppk_model = rxode2::rxode2({
    AUC(0) <- 0          #  AUC at time 0
    centr(0) <- 0        #  Central compartment
    periph(0) <- 0       #  Peripheral compartment
# Typical clearance including normalized eGFR as a covariate
    TVCl <- THETA_Cl * (eGFR / 93)^0.0043   
# Central volume of distribution with interindividual variability
    V1 <- THETA_V1 * exp(ETA_V1)
# Peripheral volume of distribution
    V2 <- THETA_V2 * exp(ETA_V2)  
# Intercompartmental clearance
    Q  <- THETA_Q  * exp(ETA_Q) 
# Individual clearance incorporating both covariate and random effect
    Cl <- TVCl     * exp(ETA_Cl)            
# Elimination rate constant
    ke  <- Cl / V1 
# Distribution rate from central to peripheral compartment
    k12 <- Q  / V1   
# Redistribution rate from peripheral to central compartment
    k21 <- Q  / V2                          
# Central compartment concentration
    Cc <- centr  / V1
# Peripheral compartment concentration
    Cp <- periph / V2                       
# Differential equation describing compartments kinetics
    d/dt(centr)  <- -ke * centr - k12 * centr + k21 * periph 
    d/dt(periph) <-  k12 * centr - k21 * periph                
    d/dt(AUC) <- Cc                       
  }),
# Combined additive and proportional residual error model
  error_model = function(f, sigma) {        
    sigma[1] + sigma[2] * f
  },
# Fixed-effect parameters
  theta = c(                                 
    THETA_Cl = 0.029,
    THETA_V1 = 6.14,
    THETA_V2 = 9.52,
    THETA_Q  = 0.026
  ),                                        
# Model covariates required for simulation and estimation
  covariates = c("eGFR"),                   
# Interindividual variability structure
  omega = lotri::lotri({                     
    ETA_Cl + ETA_V1 + ETA_V2 + ETA_Q ~
      c(0.0654, 0, 0.0255, 0, 0, 0.1213, 0, 0, 0, 0.2063)
  }),
# Residual variability parameters
  sigma = c(                                
    additive_a     = 0,
    proportional_b = 0.3392          
  )                                       
)
```

The `rxode2` model implements the following items:

-   structural population model component with the individual model (*i.e.* the model of inter-individual variability) and the covariates.
-   A function of the residual error model
-   A character vector of the covariates of the model
-   A named vector of the population estimates of the fixed effects parameters (called theta)
-   A named square variance-covariance matrix of the population parameters inter-individual variability (called omega). In general, the estimates of the variances of the random effects can be given under different parameterizations depending on the authors. In this case the coefficiente of variation ( $CV$ ) was available and the variance was computed back as $log(CV^2 + 1)$
-   The estimates of the parameters of the residual error model (called sigma)

### Model averaging and selection

When multiple models are identified for the same drug, model averaging techniques will be employed to integrate information across models. This approach acknowledges model uncertainty and allows predictions to reflect variability across plausible model structures. In the present project, model averaging will be implemented using a log-likelihood–based weighting approach, derived from the objective function values (OFV). For each model $i$ the likelihood $L_i$ is computed as: $$L_i = exp(-OFV_i/2)$$ The weights for each model are then calculated as: $$w_i = \frac{L_i}{\sum_{j=1}^{n} L_j}$$ where $n$ is the total number of models considered. These weights are subsequently used to combine model predictions, such that the final prediction for a given outcome is a weighted average of the predictions from each individual model.

The inclusion of AIC-based or alternative weighting schemes is planned as a future methodological extension, following the completion of model reconstruction and external validation activities.

### External Model Validation

External validation will be performed using retrospective clinical datasets containing TDM data for the selected antimicrobial agents. Validation will assess both population-level (a priori) and individualized (a posteriori) predictive performance of the reconstructed models. The validation pipeline will be developed as a standardized, reproducible, and user-accessible framework, enabling external users and collaborating centers to independently validate models on their own datasets.

The validation process will include the following steps:

-   **Data preparation and harmonization**: Clinical datasets will be curated to ensure data quality and consistency. Relevant patient-level covariates (e.g., demographics, body weight, renal function) will be extracted and, where necessary, transformed to match the covariate definitions and parameterizations used in the reconstructed models. Records with implausible values or insufficient information for model evaluation will be discarded.

-   **A priori (population) predictive performance:** A priori model performance will be evaluated by comparing population model predictions, generated without the use of individual concentration data, against observed TDM measurements. Predictive performance will be assessed using quantitative metrics such as the mean prediction error (MPE) to evaluate bias and the root mean square error (RMSE) to evaluate precision. Graphical diagnostics, including goodness-of-fit plots, will be used to assess agreement between predicted and observed concentration–time profiles.

-   **A posteriori (individualized) predictive performance:** A posteriori model performance will be assessed by incorporating individual TDM observations into the model to refine individual parameter estimates. Predictive performance will again be evaluated using MPE and RMSE metrics, as well as graphical diagnostics. The impact of TDM data on individual predictions will be examined to assess the clinical utility of the models for personalized dosing.

### Development of MIPD Tools

Validated models will be implemented into user-friendly Shiny applications to facilitate clinical decision support. These applications will enable clinicians to input patient-specific data and receive individualized dosing recommendations based on model simulations and target attainment analyses. Key features of the applications will include:

-   Interactive interfaces for data input and visualization of simulation results.
-   Functionality for performing Bayesian forecasting to update individual predictions based on TDM data.
-   Tools for evaluating the probability of target attainment (PTA) for various dosing regimens

### Clinical Implementation and RCT Protocol for MIPD

To ensure translational impact and clinical applicability, this project will include the development of a formal RCT protocol to evaluate the clinical implementation of MIPD-guided therapy. The trial framework will be designed to evaluate the effectiveness, safety, and clinical utility of MIPD-guided dosing strategies.

Given the multidimensional nature of antimicrobial treatment and the fact that mortality alone is an infeasible primary endpoint for evaluating dosing strategies and implementation interventions, primary and secondary endpoints will be defined using desirability-based outcome ranking frameworks, such as DOOR [@evans2015].

Depending on the clinical syndrome under investigation, primary and secondary endpoints will be defined using desirability-based outcome ranking frameworks, such as DOOR, informed by existing methodological literature or adapted to the specific clinical context of antimicrobial therapy [@ong2023].

Adaptive trial components will be explored, with particular emphasis on population enrichment strategies. This will include adaptive recruitment frameworks that progressively focus on patient subgroups most likely to benefit from MIPD-guided therapy, such as individuals with high PK variability, altered physiology, or organ dysfunction [@Pallmann2018].
