---
title: "Dalbavacin report"
output: html_document
date: "`r Sys.Date()`"
params:
   plot_dalba: NULL,
   combined_plot: NULL,
   doses: NULL,
   patient_tb: NULL,
   date_to_display: NULL, 
   days_from_last_dose: NULL
   first_below_target_ab: NULL
   free_drug: NULL
   target_mic: NULL
   target_auc: NULL
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
params$combined_plot
```


```{r echo=FALSE, message=FALSE, warning=FALSE}
params$plot_dalba
```

For the estimation the target AUC/MIC selected was `r params$target_auc` mg·h/L, with and a target MIC of `r params$target_mic` and a free drug estimation of `r params$free_drug * 100`%.

According to the model averaging dalabavacin should be redose on the `r params$date_to_display` (`r params$days_from_last_dose` days from the last given dose) when the last fAUC0-24/MIC is estimated to be `r params$first_below_target_ab$AUC24`  mg·h/L with a corresponding Cmin of `r round(params$first_below_target_ab$Cc,2)` mg/L. 


```{r echo=FALSE, message=FALSE, warning=FALSE}
knitr::kable(params$patient_tb, caption = "Patient Parameters", align = "c")
```

